---
title: "3.2 - Time series componenets"
format: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(paged.print = FALSE)
```

```{r message=FALSE}
library(fpp3)
```

-   *Additive*: $y_{t} = S_{t} + T_{t} + R_t$
-   *Multiplicative*: $y_{t} = S_{t} \times T_{t} \times R_t$

To can use a log transformation and then additive to achieve multiplicative.

## Example: US retail employment

We will decompose the number of persons employed in retail.

```{r}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
autoplot(us_retail_employment, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

```{r}
dcmp <- us_retail_employment |>
  model(stl = STL(Employed))
components(dcmp)
```

```{r}
components(dcmp) |>
  as_tsibble() |>
  autoplot(Employed, colour="gray") +
  geom_line(aes(y=trend), color="#D55E00") +
  labs(
    y = "Persons (thousands)",
    title = "Total employment in US retail"
  )
```

```{r}
components(dcmp) |> autoplot()
```

## Seasonally adjusted data

```{r}
components(dcmp) |>
  as_tsibble() |>
  autoplot(Employed, color='gray') +
  geom_line(aes(y=season_adjust), colour="#0072B2") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  
```

Seasonally adjusted series contain the remainder component as well as the trend-cycle. Therefore, they are not “smooth”, and “downturns” or “upturns” can be misleading. If the purpose is to look for turning points in a series, and interpret any changes in direction, then it is better to use the trend-cycle component rather than the seasonally adjusted data.

