---
title: "3.4 - Moving Averages"
format: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(paged.print = FALSE)
```

```{r message=FALSE}
library(fpp3)
```


## MA Smoothing

$$
\begin{equation}
  \hat{T}_{t} = \frac{1}{m} \sum_{j=-k}^k y_{t+j}
\end{equation}
$$

where $m=2k+1$.

```{r}
global_economy |>
  filter(Country == "Australia") |>
  autoplot(Exports) +
  labs(y="% of GDP", title = "Total Australian exports")
```

```{r}
aus_exports <- global_economy |>
  filter(Country == "Australia") |>
  mutate(
    `5-MA` = slider::slide_dbl(Exports, mean,
                               .before = 2,
                               .after = 2,
                               .complete = TRUE)
  )
```

```{r}
aus_exports |>
  autoplot(Exports) +
  geom_line(aes(y = `5-MA`), color = "#D55E00") +
  labs(y = "% of GDP",
       title = "Total Australian exports")
```

## Moving averages of moving averages

For even-order moving averages, eg. periods of 12 or 4. Apply an M2 period after the even selection.

```{r}
beer <- aus_production |>
  filter(year(Quarter) >= 1992) |>
  select(Quarter, Beer)
beer_ma <- beer |>
  mutate(
    `4-MA` = slider::slide_dbl(Beer, mean,
                               .before = 1, .after = 2,
                               .complete = TRUE),
    `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                                 .before = 1, .after = 0,
                                 .complete = TRUE)
  )
```

```{r}
beer_ma
```

## Estimate trend-cycle with seasonal data

```{r}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
```

```{r}
us_retail_employment_ma <- us_retail_employment |>
  mutate(
    `12-MA` = slider::slide_dbl(Employed, mean,
                                .before = 5, .after = 6,
                                .complete = TRUE),
    `2x12-MA` = slider::slide_dbl(`12-MA`, mean,
                                  .before = 1, .after = 0,
                                  .complete = TRUE)
  )
us_retail_employment_ma |>
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = `2x12-MA`), colour = "#D55E00") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

### Weighted moving averages

Combinations of moving averages result in weighted moving averages. For example, the $2×4$-MA discussed above is equivalent to a weighted 5-MA with weights given by $\left[\frac{1}{8},\frac{1}{4},\frac{1}{4},\frac{1}{4},\frac{1}{8}\right]$. In general, a weighted m-MA can be written as $$\hat{T}_t = \sum_{j=-k}^k a_j y_{t+j}$$, where k=(m−1)/2, and the weights are given by $[a_{-k}, \dots, a_k$. It is important that the weights all sum to one and that they are symmetric so that $a_j = a_{-j}$. The simple m-MA is a special case where all of the weights are equal to 1/m.

A major advantage of weighted moving averages is that they yield a smoother estimate of the trend-cycle. Instead of observations entering and leaving the calculation at full weight, their weights slowly increase and then slowly decrease, resulting in a smoother curve.

