---
title: "3.6 - STL Decomposition"
format: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(paged.print = FALSE)
```

```{r}
library(fpp3)
```

# 3.6 STL decomposition

STL has several advantages over classical decomposition, and the SEATS and X-11 methods:

-   Unlike SEATS and X-11, STL will handle any type of seasonality, not only monthly and quarterly data.
-   The seasonal component is allowed to change over time, and the rate of change can be controlled by the user.
-   The smoothness of the trend-cycle can also be controlled by the user.
-   It can be robust to outliers (i.e., the user can specify a robust decomposition), so that occasional unusual observations will not affect the estimates of the trend-cycle and seasonal components. They will, however, affect the remainder component.

On the other hand, STL has some disadvantages. In particular, it does not handle trading day or calendar variation automatically, and it only provides facilities for additive decompositions.

```{r}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
```

```{r}
us_retail_employment |>
  model(
    STL(Employed ~ trend(window = 7) +
                   season(window = "periodic"),
    robust = TRUE)) |>
  components() |>
  autoplot()
```
